CodeBook
========================================================
This is a code book that describes the variables, the data, and any transformations or work that was performed to clean up the data and generate tidy.txt file


```{r}
source("run_analysis.R")
```

## Variable list and description
This are the first 12 columns of the final dataset:
```{r}
names(tidy)[1:10]
```

| Tables   |      Description      | 
|:----------|:-------------|
| subjectId |  Subject identifier. Its range is from 1 to 30. |
| activityName | Activity performed by individuals e.g. Walking, Sleeping etc | 
| Measures | There are 66 observations in the set. The measures represent data that came from accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. Example, timeBodyAccelerator-mean()-X ,timeGravityAccelerator-mean()-Y| 

## Dataset Structure
There are 66 variables across 30 participants and 6 activities. Not all participants have measures for each activity.
```{r}
str(tidy)
```

Data sample
```{r}
head(tidy)
```

## Data transformation steps
* Load dplyr package by using library(dplyr) command
* Loading x_train and x_test data sets
```{r}
x_train<-read.table(file="UCI HAR Dataset/train/X_train.txt")
```
* Loading metadata e.g. features, activities etc.
```{r}
act_labels <-read.table(file="UCI HAR Dataset/activity_labels.txt")
features <-read.table(file="UCI HAR Dataset/features.txt")
features <-features[,2]
features_char<-as.character(features)
```
* Replacing numeric activity ids with names  e.g. Walking, Staying etc
```{r}
train_labels<-merge(y_train,act_labels)
test_labels<-merge(y_test,act_labels)
```
* Binding data sets (both train and test) with corresponding subject data
```{r}
train<-cbind(subject_train,train_labels$V2,x_train)
test<-cbind(subject_test,test_labels$V2,x_test)
```
* Union train and test data sets into a single set
```{r}
merged<-rbind(train,test)
```
* Using RegEx selecting columns that contain mean() or std() in their names. It reduces the set from 561 observations to 66.
```{r}
merged<-subset(merged, select = grep("mean\\(\\)|std\\(\\)|subjectId|activityName", names(merged),value=T))
```
* Replace selected column names with more meaningful onces e.g. Accelerator instead of Acc, time instead of t. Below is an example of the substring replacement for Acc.
```{r}
names(merged) <- gsub("Acc", "Accelerator", names(merged))
```
* Grouping data in a set by subjectId and activityName and aggregating values  using mean function.
```{r}
tidy<-merged %>% group_by(subjectId, activityName) %>% summarise_each(funs(mean))
```
* Publishing tidy.txt file
```{r}
write.table(tidy, file = "Tidy.txt", quote = FALSE,sep = ",",row.names = FALSE)
```

