CodeBook
========================================================
This is a code book that describes the variables, the data, and any transformations or work that was performed to clean up the data and generate tidy.txt file

For a full code description and full data transformation steps please see run_analysis.R file. 

## Variable list and description


| Field   |      Description      | 
|:----------|:-------------|
| subjectId |  Subject identifier. Its range is from 1 to 30. |
| activityName | Activity performed by individuals e.g. Walking, Sleeping etc | 
| timeBodyAccelerator-mean()-X | There are 66 observations in the set. The measures represent data that came from accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. | 
| timeBodyAccelerator-mean()-Y ||
| timeBodyAccelerator-mean()-Z ||
| ... ||
| timeGravityAccelerator-mean()-Z ||


## Dataset Structure
There are 66 variables across 30 participants and 6 activities e.g. Walking, Sleeping etc. 
```{r}
str(tidy)
```
This is a part of the str(tidy) result:

Classes ‘grouped_df’, ‘tbl_df’, ‘tbl’ and 'data.frame':	40 obs. of  68 variables:
| Tables   |      Description      | 
|:----------|:-------------|
|$ subjectId| int  1 2 3 4 4 5 6 6 7 8 ...|
|$ activityName                          |  Factor w/ 6 levels "LAYING","SITTING",..: 4 4 4 4 6 4 4 6 6 6 ...|
 |$ timeBodyAccelerator-mean()-X          |  num  0.266 0.273 0.273 0.277 0.27 ...|
| $ timeBodyAccelerator-mean()-Y          |  num  -0.0183 -0.0191 -0.0179 -0.0133 -0.0171 ...|
 |$ timeBodyAccelerator-mean()-Z          |  num  -0.108 -0.116 -0.106 -0.106 -0.11 ...|
| $ timeBodyAccelerator-std()-X           |  num  -0.546 -0.606 -0.623 -0.684 -0.481 ...|
| $ timeBodyAccelerator-std()-Y           |  num  -0.368 -0.429 -0.48 -0.59 -0.384 ..|
 |
## Data sample
```{r}
head(tidy)
```
Source: local data frame [6 x 68]
Groups: subjectId [5]

  subjectId     activityName timeBodyAccelerator-mean()-X timeBodyAccelerator-mean()-Y timeBodyAccelerator-mean()-Z
      (int)           (fctr)                        (dbl)                        (dbl)                        (dbl)
1         1          WALKING                    0.2656969                  -0.01829817                   -0.1078457
2         2          WALKING                    0.2731131                  -0.01913232                   -0.1156500
3         3          WALKING                    0.2734287                  -0.01785607                   -0.1064926
4         4          WALKING                    0.2770345                  -0.01334968                   -0.1059161
5         4 WALKING_UPSTAIRS                    0.2696859                  -0.01710851                   -0.1100534

## Data transformation steps
* Load dplyr package by using library(dplyr) command
* Loading x_train and x_test data sets
```{r}
x_train<-read.table(file="UCI HAR Dataset/train/X_train.txt")
```
* Loading metadata e.g. features, activities etc.
```{r}
act_labels <-read.table(file="UCI HAR Dataset/activity_labels.txt")
features <-read.table(file="UCI HAR Dataset/features.txt")
features <-features[,2]
features_char<-as.character(features)
```
* Replacing numeric activity ids with names  e.g. Walking, Staying etc
```{r}
train_labels<-merge(y_train,act_labels)
test_labels<-merge(y_test,act_labels)
```
* Binding data sets (both train and test) with corresponding subject data
```{r}
train<-cbind(subject_train,train_labels$V2,x_train)
test<-cbind(subject_test,test_labels$V2,x_test)
```
* Union train and test data sets into a single set
```{r}
merged<-rbind(train,test)
```
* Using RegEx selecting columns that contain mean() or std() in their names. It reduces the set from 561 observations to 66.
```{r}
merged<-subset(merged, select = grep("mean\\(\\)|std\\(\\)|subjectId|activityName", names(merged),value=T))
```
* Replace selected column names with more meaningful onces e.g. Accelerator instead of Acc, time instead of t. Below is an example of the substring replacement for Acc.
```{r}
names(merged) <- gsub("Acc", "Accelerator", names(merged))
```
* Grouping data in a set by subjectId and activityName and aggregating values  using mean function.
```{r}
tidy<-merged %>% group_by(subjectId, activityName) %>% summarise_each(funs(mean))
```
* Publishing tidy.txt file
```{r}
write.table(tidy, file = "Tidy.txt", quote = FALSE,sep = ",",row.names = FALSE)
```

