CodeBook
========================================================
This is a code book that describes the variables, the data, and any transformations or work that was performed to clean up the data and generate tidy.txt file

For a full code description and full data transformation steps please see run_analysis.R file. 

## Variable list and description


| Field   |      Description      | 
|:----------|:-------------|
| subjectId |  Subject identifier. Its range is from 1 to 30. |
| activityName | Activity performed by individuals e.g. Walking, Sleeping etc | 
| timeBodyAccelerator-mean()-X | There are 66 observations in the set. The measures represent data that came from accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. | 
| timeBodyAccelerator-mean()-Y ||
| timeBodyAccelerator-mean()-Z ||
| ... ||
| timeGravityAccelerator-mean()-Z ||


## Dataset Structure
There are 66 variables across 30 participants and 6 activities e.g. Walking, Sleeping etc. 
```{r}
str(tidy)
```
This is a part of the str(tidy) result:

Classes ‘grouped_df’, ‘tbl_df’, ‘tbl’ and 'data.frame':	40 obs. of  68 variables:

| Field   |      Structure Details      | 
|:----------|:-------------|
|$ subjectId| int  1 2 3 4 4 5 6 6 7 8 ...|
|$ activityName                          |  Factor w/ 6 levels "LAYING","SITTING",..: 4 4 4 4 6 4 4 6 6 6 ...|
|$ timeBodyAccelerator-mean()-X          |  num  0.266 0.273 0.273 0.277 0.27 ...|
| $ timeBodyAccelerator-mean()-Y          |  num  -0.0183 -0.0191 -0.0179 -0.0133 -0.0171 ...|
|$ timeBodyAccelerator-mean()-Z          |  num  -0.108 -0.116 -0.106 -0.106 -0.11 ...|
| $ timeBodyAccelerator-std()-X           |  num  -0.546 -0.606 -0.623 -0.684 -0.481 ...|
| $ timeBodyAccelerator-std()-Y           |  num  -0.368 -0.429 -0.48 -0.59 -0.384 ..|
 
## Data sample
```{r}
head(tidy)
```
Source: local data frame [6 x 68]
Groups: subjectId [5]

| subjectId   |      activityName      | timeBodyAccelerator-mean()-X |timeBodyAccelerator-mean()-Y|timeBodyAccelerator-mean()-Z|...|
|:----------|:-------------|:-------------|:-------------|:-------------|:-------------|
|      (int)|           (fctr) |                       (dbl) |                       (dbl) |                       (dbl)|...|
|        1 |            LAYING |           0.2215982 |                -0.040513953|                   -0.1132036||
|        1 |           SITTING  |          0.2612376 |                -0.001308288 |                  -0.1045442||
|       1 |          STANDING  |          0.2789176 |                -0.016137590 |                  -0.1106018||
|        1 |           WALKING    |        0.2773308 |                -0.017383819 |                  -0.1111481||
|        1 |WALKING_DOWNSTAIRS   |         0.2891883 |                -0.009918505  |                 -0.1075662||
|        1 |  WALKING_UPSTAIRS   |         0.2554617 |                -0.023953149  |                 -0.0973020||
## Data transformation steps
* Load dplyr package by using library(dplyr) command
* Loading x_train and x_test data sets. Full code is in the run_analysis.R.
```{r}
x_train<-read.table(file="UCI HAR Dataset/train/X_train.txt")
```
* Loading metadata e.g. features, activities etc.
```{r}
act_labels <-read.table(file="UCI HAR Dataset/activity_labels.txt")
features <-read.table(file="UCI HAR Dataset/features.txt")
features <-features[,2]
features_char<-as.character(features)
```

* Bind together main sets (x_train and x_test) with its labels and subjects
```{r}
train<-cbind(subjectId=subject_train$V1, activityId=y_train$V1,x_train )
test<-cbind(subjectId=subject_test$V1, activityId=y_test$V1,x_test)
```
* Now union the test and train sets 
```{r}
merged<-rbind(train,test)
```
* Add descriptions of the activities to enrich the data
```{r}
merged<-merge(act_labels,merged,  by="activityId")
```

* Clean up the merged dataset by removing redundant columns
```{r}
merged<-merged[ , !(names(merged) %in% c("activityId", "activityName.y"))]
```

* Rename activiyName.x column to activityName
```{r}
names(merged)[1] <- "activityName"
```
* Name the columns of the merged set 
```{r}
names(merged)[3:563]<-features_char[1:561]
```

* Extract only required columns plus subjectId and activityName
```{r}
merged<-subset(merged, select = grep("mean\\(\\)|std\\(\\)|subjectId|activityName", names(merged),value=T))
```
* Appropriately labels the data set with descriptive variable names
```{r}
names(merged) <- gsub("Acc", "Accelerator", names(merged))
```

...


* Generate an independent tidy data set with the average of each variable  for each activity and each subject (e.g. person).
```{r}
tidy<-merged %>% group_by(subjectId, activityName) %>% summarise_each(funs(mean))
write.table(tidy, file = "tidy.txt", quote = FALSE,sep = ",",row.names = FALSE)

```



